Further Evidence for Modified Spin-down in Sun-like Stars:\\Pile-ups in the Temperature--Period Distribution

### 2021-12-17 src/ms.tex
jvs comments

Notably, both models fail to reproduce the observed short-period pile-up, possibly due to the assumption of solid-body rotation in both models. At early times, Sun-like and low-mass stars are expected to have strong radial differential rotation due to their rapid collapse onto the main-sequence. The core and envelope at these times are thus assumed to be decoupled. However, the core and envelope are expected to couple on timescales of a few tens of million years for Sun-like stars \citep{Denissenkov2010, GalletBouvier2015, Lanzafame2015} or hundreds of million years for low-mass stars \citep{GalletBouvier2015,Lanzafame2015,Somers2016}. When the core and envelope couple, the star experiences a spin-up torque such that so-called ``two-zone'' models \citep{MacGregor1991} spin more rapidly than solid-body rotators at the same age. This hypothesis is discussed further in \S\ref{subsec:shortperiod}.

\begin{figure*}
    \centering 
    \includegraphics[width=0.49\linewidth]{./figures/std-model-cks.pdf}
    \includegraphics[width=0.49\linewidth]{./figures/wmb-model-cks.pdf}
    \includegraphics[width=0.49\linewidth]{./figures/std-model-lamost.pdf}
    \includegraphics[width=0.49\linewidth]{./figures/wmb-model-lamost.pdf}
    \caption{The \teff-\prot plane for the CKS sample (top panels) and LAMOST sample (bottom panels) in comparison to the standard and WMB models presented in \citet{vanSaders2019}. The excess of stars with periods $\gtrsim$~35 days, seen most clearly in the models, are due to subgiants which have expanded off the main-sequence \citep[referred to as the long-period overdensity in][]{vanSaders2019}.}
    \label{fig:models}
\end{figure*}

>>>>

Notably, both models fail to reproduce the observed short-period pile-up, possibly due to the assumption of solid-body rotation in both models. At early times, Sun-like and low-mass stars are expected to have strong radial differential rotation due to their rapid collapse onto the main-sequence. The core and envelope at these times are thus assumed to be decoupled. However, the core and envelope are expected to couple on timescales of a few tens of million years for Sun-like stars \citep{Denissenkov2010, GalletBouvier2015, Lanzafame2015} or hundreds of million years for low-mass stars \citep{GalletBouvier2015,Lanzafame2015,Somers2016}. When this happens, angular momentum can be transferred from the core to the envelope at a rate comparable to the rate at which angular momentum is lost via magnetized winds. Consequently, so-called ``two-zone'' models \citep{MacGregor1991} spin more rapidly than solid-body rotators at the same age. This hypothesis is discussed further in \S\ref{subsec:shortperiod}.

\begin{figure*}
    \centering 
    \includegraphics[width=0.49\linewidth]{./figures/std-model-cks.pdf}
    \includegraphics[width=0.49\linewidth]{./figures/wmb-model-cks.pdf}
    \includegraphics[width=0.49\linewidth]{./figures/std-model-lamost.pdf}
    \includegraphics[width=0.49\linewidth]{./figures/wmb-model-lamost.pdf}
    \caption{The \teff-\prot plane for the CKS sample (top panels) and LAMOST sample (bottom panels) in comparison to the standard and WMB models presented in \citet{vanSaders2019}. The excess of stars with periods $\gtrsim$~35 days, seen most clearly in the models, are due to subgiants which have expanded off the main-sequence \citep[referred to as the long-period overdensity in][]{vanSaders2019}. Rotation periods for the CKS sample here are sourced from the \citet{David2021} compilation.}
    \label{fig:models}
\end{figure*}
